Finiteness in Cubical Type Theory

### 2020-06-16 sections/introduction.tex
more agda conversion

Finally, we will also want to optimise our solutions and solver: for this we
will remove equivalent expressions, which can be accomplished with quotients.
\section{Notation and Background}
We work in Cubical Type Theory \cite{cohenCubicalTypeTheory2016}.
The notation we use comes from the conventions used in the Agda standard
library.
\begin{definition}[Type]
  We use \(\AgdaDatatype{Type}\) to denote the universe of (small) types.
  The universe level is denoted with a subscript number, starting at 0.
  ``Type families'' are functions into \(\AgdaDatatype{Type}\).
\end{definition}
\begin{definition}[Basic Types]
  The three basic types---often called 0, 1, and 2 in MLTT---here will be
  denoted with their more common names: \(\bot\), \(\top\), and
  \(\mathbf{Bool}\), respectively.
  \begin{multicols}{3} \centering
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{bot}
    \end{agdalisting}

@@

  \begin{multicols}{3} \centering
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{pi-syntax-1}
    \end{agdalisting}

@@

\begin{description}
  \item[Disjoint Union] We define disjoint union as an inductive type.
    \begin{equation}
      \begin{alignedat}{3}
        A \uplus B \coloneqq & \;
        \text{inl} &: A \rightarrow A \uplus B ; \\
        | & \;  \text{inr} &: B \rightarrow A \uplus B ;
      \end{alignedat}
    \end{equation}
    It is also expressible with only \(\Sigma\):
    \(A \uplus B \simeq \Sigma(x : \mathbf{Bool}) , \text{if } x \text{ then
    } A \text{ else } B \).
  \item[Equalities, equivalences, and paths] We use the symbol \(\coloneqq\)
    for definitions.
    \(\simeq\) will be used for equivalences, and \(\equiv\) for equalities.
    Of course, we know that \((A \simeq B) \simeq (A \equiv B)\) by univalence,
    so the distinction isn't terribly important in usage: we will only use one
    or the other as a suggestion of how we constructed it or how it is to be
    used.
  \item[Lambdas] \todo{Figure out how to describe lambdas}
\end{description}
\subsection{Cubical Type Theory}
Cubical Type Theory \cite{cohenCubicalTypeTheory2016} is a constructive type
theory with an implementation in Cubical Agda
\cite{vezzosiCubicalAgdaDependently2019}.
It allows us to do much of the same theory as in HoTT, but crucially the
univalence ``axiom'' is a \emph{theorem}, giving it computational content.
\begin{definition}[Path Types] \label{path-types}
  The equality type (which we denote with \(\equiv\)) in CuTT is the type of
  Paths\footnotemark.
  The internal structure of paths is largely irrelevant to us here, as we will
  generally treat \(\equiv\) as a black-box equivalence relation with
  substitution and congruence.
\end{definition}

>>>>

Finally, we will also want to optimise our solutions and solver: for this we
will remove equivalent expressions, which can be accomplished with quotients.
\section{Notation and Background}
We work in Cubical Type Theory \cite{cohenCubicalTypeTheory2016}, specifically
Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}.
Cubical Agda is a dependently-typed functional programming language, based on
Martin-Löf Intuitionistic Type Theory, with a Haskell-like syntax.

Being a dependently-typed language, we'll have to be clear about what we mean
when we say ``type'' in Agda.
\begin{definition}[Type]
  We use \(\AgdaDatatype{Type}\) to denote the universe of (small) types.
  The universe level is denoted with a subscript number, starting at 0.
  ``Type families'' are functions into \(\AgdaDatatype{Type}\).
\end{definition}

The are two broad ways to define types in Agda: as an inductive
\(\AgdaKeyword{data}\) type, similar to data type definitions in Haskell, or as
a \(\AgdaKeyword{record}\).
Here we'll define the basic type formers used in MLTT.
\begin{definition}[Basic Types]
  The three basic types---often called 0, 1, and 2 in MLTT---here will be
  denoted with their more common names: \(\bot\), \(\top\), and
  \(\mathbf{Bool}\), respectively.
  \begin{multicols}{3} \centering
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{bot}
    \end{agdalisting}

@@

  \begin{multicols}{3}
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{pi-syntax-1}
    \end{agdalisting}

@@

\begin{definition}[Disjoint Union]
  We define disjoint union as an inductive type.
  \begin{agdalisting}
    \ExecuteMetaData[agda/Snippets/Introduction.tex]{disj-union}
  \end{agdalisting}
  It is also expressible with only \(\Sigma\):
  \begin{agdalisting}
    \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-disj-union}
  \end{agdalisting}
\end{definition}
\begin{definition}[Equalities, equivalences, and paths] We use the symbol \(\coloneqq\)
  for definitions.
  \(\simeq\) will be used for equivalences, and \(\equiv\) for equalities.
  Of course, we know that \((A \simeq B) \simeq (A \equiv B)\) by univalence,
  so the distinction isn't terribly important in usage: we will only use one
  or the other as a suggestion of how we constructed it or how it is to be
  used.
\end{definition}

\begin{definition}[Path Types] \label{path-types}
  The equality type (which we denote with \(\equiv\)) in CuTT is the type of
  Paths\footnotemark.
  The internal structure of paths is largely irrelevant to us here, as we will
  generally treat \(\equiv\) as a black-box equivalence relation with
  substitution and congruence.
\end{definition}
